---
title: "Job Proficiency Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Job Proficiency Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette analyzes job proficiency data using a dataset stored in `data.xlsx`. We visualize the relationship between job proficiency and test scores using scatterplots.

## Load Data

```{r load-data}
library(readxl)

# Read data from "data.xlsx" without headers
jobs <- read_excel("data.xlsx", col_names = FALSE)

# Assign column names to the dataframe
colnames(jobs) <- c("proficiency", "t1", "t2", "t3", "t4")

# Display data summary
summary(jobs)
```

## Visualize Data

We create scatterplots to visualize the relationship between job proficiency and each test score.

```{r scatterplots}
# Set appropriate plot size
options(repr.plot.width = 8, repr.plot.height = 8)

# Set plotting area to 2x2 grid
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1)) # Add margin settings

# Scatterplot for proficiency vs t1
plot(jobs$t1, jobs$proficiency,
     xlab = "Test 1 (t1)",
     ylab = "Job Proficiency",
     main = "Proficiency vs t1")

# Scatterplot for proficiency vs t2
plot(jobs$t2, jobs$proficiency,
     xlab = "Test 2 (t2)",
     ylab = "Job Proficiency",
     main = "Proficiency vs t2")

# Scatterplot for proficiency vs t3
plot(jobs$t3, jobs$proficiency,
     xlab = "Test 3 (t3)",
     ylab = "Job Proficiency",
     main = "Proficiency vs t3")

# Scatterplot for proficiency vs t4
plot(jobs$t4, jobs$proficiency,
     xlab = "Test 4 (t4)",
     ylab = "Job Proficiency",
     main = "Proficiency vs t4")

# Reset plotting area to default (1x1)
par(mfrow = c(1, 1))
```

## Conclusion

The scatterplots provide insights into how each test score correlates with job proficiency. Further statistical analysis could quantify these relationships.
